Industrial Organization III: \\ Final Paper

### 2024-09-05 draft/draft.tex
Fixed UV decomposition

MovieLens produces a public-use movie review dataset; it contains more than 25 million (CHECK THIS NUMBER) 1-to-5 star reviews of movies by users. Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions. Thus, for any given pair of movies, I can compute similarity under both the text-based and collaborative filtering methods. If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.

@@

For my collaborative filtering algorithm, I use UV-decomposition for matrix completion as described in Leskovec (CITE HERE). Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews. However, the matrix is generally sparse; most users have not reviewed most movies. UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV^T$ approximates the original matrix. This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches. By minimizing the sum of squared errors among projections in the \emph{observed} entries of the matrix, we hope to project the \emph{unobserved} entries of the matrix as well.

@@

\subsection{Validation}

@@

\subsection{Comparison}

LOREM IPSUM

>>>>

MovieLens produces a public-use movie review dataset; it contains more than 25 million (CHECK THIS NUMBER) 1-to-5 star reviews of movies by users. Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions. Thus, for any given pair of movies, I can compute similarity under both the text-based and collaborative filtering methods. If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental. I validate my text-based model by comparing it against two collaborative filtering algorithms: topic-specific PageRank and UV decomposition.

\subsection{Topic-Specific PageRank}

Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of movies, as implemented by CITE FAST PAGERANK HERE. PageRank is an algorithm for detecting the most important nodes in a network; it was originally developed by Google to rank webpages. The algorithm works by randomly walking through the network, with a probability of jumping to a random node at each step. The importance of a node is the proportion of time the random walker spends at that node. TSPR is an extension of the algorithm which always ``jumps'' to a specific set of nodes rather than a random node.

For my application, I implement TSPR as follows. To find movies similar to specific movie $x$,:
\begin{itemize}
    \item The algorithm randomly walks over the bipartite network of users and movies. 
    \item At each step, the algorithm has a 15\% chance of jumping back to $x$
    \item If the algorithm does not jump to $x$, it takes a random step to a neighbor of its current node
    \item Random draws are weighted by the number of stars in the review; e.g., a ``5-star'' review is five times as likely to be drawn as a ``1-star'' review
\end{itemize}

Running this algorithm until convergence provides a measure of how central any given movie $y$ is relative to movie $x$. However, we are still not quite done, as some movies are just \emph{inherently} central; if everyone likes \emph{Forrest Gump,} then it's not particularly informative that fans of movie $x$ \emph{also} like \emph{Forrest Gump}. Thus, I divide the TSPR score of $y$ by $y$'s \emph{non}-topic specific PageRank score. This gives a measure of how much more central $y$ is to $x$ than it is to the average movie. Given this is a ratio, it has propensity to explode if $y$'s centrality is very small; thus, I prefer to take the log of this ratio for my analysis.

Once I have each movie's TSPR score for each other movie, I can compare these scores to the cosine similarity of the text embeddings. Note that the TSPR scores are not symmetric; that is, the TSPR score of movie $x$ to movie $y$ is not necessarily the same as the TSPR score of movie $y$ to movie $x$. Thus, for each pair of movies, I actually plot \emph{two} measures of similarity; one from $x$ to $y$ and another from $y$ to $x$; these pairs have the same text-based cosine similarity, but different TSPR values.

A common weakness of collaborative filtering algorithms like TSPR is that they require significant data before they become effective. Thus, I compute TSPR values only for movies which receive at least 10,000 reviews in the MovieLens data. This leaves me with 258 movies, or 66,564 pairs of movies. To assess the importance of this margin, I then plot the coefficient of regressing TSPR on text similarity for thresholds 10,000 to 50,000. I also plot the number of movie pairs included in the sample for each threshold value, to assess the trade-off of ``more ratings per movie'' vs. ``more movies.'' The results of this plot are visible in figure \ref{fig:tspr_threshold}.

\begin{figure}
    \begin{center}
    \includegraphics[width=\textwidth]{../movielens/output/threshold_vs_coefficient.png}
    \caption{Coefficient of log(TSPR) on Text Similarity by Threshold}
    \label{fig:tspr_threshold}
    \end{center}
\end{figure}

This plot shows that the coefficient on text similarity for predicting TSPR is consistently and significantly positive for nearly all thresholds. Thus, the content of the text embeddings \emph{does} seem to have predictive power for which movies are similar based on revealed preference; that is, the patterns I see in the text-based model are not coincidental and reflect consumers' true preferences.

Moreover, while the number of pairs included remains large enough to have useful confidence intervals, the coefficient is increasing in the threshold value. If we think TSPR increases in quality with data, this suggests that as TSPR becomes higher quality, it looks increasingly like our text-based cosine similarity measure. This exercise highlights an additional strength of my cosine similarity measure; while collaborative filtering needs tens of thousands of reviews per movie to generate useful predictions, the embeddings-based method can generate useful predictions with only a few sentences of text.

@@

I implement UV-decomposition for matrix completion as described in Leskovec (CITE HERE). Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews. However, the matrix is generally sparse; most users have not reviewed most movies. UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV^T$ approximates the original matrix. This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches. By minimizing the sum of squared errors among projections in the \emph{observed} entries of the matrix, we hope to project the \emph{unobserved} entries of the matrix as well.

@@

@@

I find XXXXXX
